Efficiently summarizing relationships in large samples: a general duality between statistics of genealogies and genomes

### 2020-01-18 treestats_paper.tex
fig 1 draft

\begin{figure}
    \begin{center}
    \includegraphics{figures/divergence_diagram}
    \end{center}
    \caption{
        An example of computing sequence divergence between red nodes and blue nodes
        from seven mutations at distinct sites in a sequence of length 10 using the tree.
        The summary function, $f(a,b) = (a/3)(1-b/2)$ gives the probability that
        an allele found in $a$ of the red nodes and $b$ of the blue nodes is present in
        a randomly chosen red node but not in a randomly chosen blue node.
        \textbf{(a)} Weights are assigned to each subtree by adding together the weights of any children,
        plus the weight of the node itself, if it is a sample.
        \textbf{(b)} In this example, subtree weights $(a,b)$ record how many red nodes ($a$)
        and how many blue nodes ($b$) lie in that subtree.
        \textbf{(c)} The summary function shown is applied to both the weight of the subtree
        and the remaining weight of each mutation,
        so for instance the mutation above the node with weights $(2,1)$
        contributes $f(2,1) + f(1,1) = 1/3 + 1/6 = 1/2$.
        The final value for divergence is the sum of these values divided by 10,
        the total number of sites.
        \label{fig:divergence_diagram}
    }
\end{figure}

>>>>

\begin{figure}
    \begin{center}
    \includegraphics{figures/divergence_diagram_simple}
    \end{center}
    \caption{
        \textbf{FIXME:}
        An example of computing sequence divergence between red nodes and blue nodes
        from seven mutations at distinct sites in a sequence of length 10 using the tree.
        The summary function, $f(a,b) = (a/3)(1-b/2)$ gives the probability that
        an allele found in $a$ of the red nodes and $b$ of the blue nodes is present in
        a randomly chosen red node but not in a randomly chosen blue node.
        \textbf{(a)} Weights are assigned to each subtree by adding together the weights of any children,
        plus the weight of the node itself, if it is a sample.
        \textbf{(b)} In this example, subtree weights $(a,b)$ record how many red nodes ($a$)
        and how many blue nodes ($b$) lie in that subtree.
        \textbf{(c)} The summary function shown is applied to both the weight of the subtree
        and the remaining weight of each mutation,
        so for instance the mutation above the node with weights $(2,1)$
        contributes $f(2,1) + f(1,1) = 1/3 + 1/6 = 1/2$.
        The final value for divergence is the sum of these values divided by 10,
        the total number of sites.
        \label{fig:divergence_diagram}
    }
\end{figure}
